\sys: Light-touch Data-Parallel Shell Processing

### 2020-05-23 pldi/draft.tex
Minor edits and adding results in evaluation

\begin{table*}[t]
\center
\footnotesize
% \setlength\tabcolsep{3pt}
\caption{
  \footnotesize{
    \textbf{Summary of shell one-liners}.
    Small pipelines (\ie ones usually developed on-the-fly) drawn from various sources and applied to large datasets.
    % Names correspond to the speedup plots (Fig.~\ref{fig:microbenchmarks}).
    Structure summarizes the different classes of commands used in the script.
    Input and seq. time report on the input size fed to the script and the timing of its sequential execution.
    Nodes and compile time report on \sys's resulting \todo{AST} size and compilation time for two indicative parallelization configurations.
    % Highlights notes a few key features of each script.
  }
}
\input{microbenchmarks-table}
\label{tab:eval}
\end{table*}

@@

\heading{Programs}
%>\kk{How can we include a result for grep-light?  This is the IO
%>  intensive CPU-light grep. \sys has no speedup but also no slowdown
%>  (even if it IO heavy). Should we just mention that in text in
%>  passing?}
%>
%>\kk{Remember to mention that for spell and bigrams we dont show other
%>  parallel because they are all close to 1 (no speedup).}
Tab.~\ref{tab:eval} summarizes the collection of programs shown in Fig.~\ref{fig:microbenchmarks}, which in turn reports on their performance characteristics.
Grep and Grep-light are in \sta, with the former centered around an expensive DFA-based backtracking expression.
% a DFA-based regular expression matching phase;
 % while the \unix \ttt{grep} command defaults to a Thompson NFA, this particular expression makes use of DFA-based backtracking patterns resulting in high runtime overhead.
Sort is a short script centered around a \pur command.
Wf and Top-n are based on McIlroy's classic word-counting program~\cite{bentley1986literate};
  they use sorting, rather than tabulation, to identify high-frequency terms in a corpus.
% \tr{The \ttt{bi-gram} pipeline calculates n-grams of base two;
%   it makes clever use of \ttt{tail} to shift a stream by one word and \ttt{paste} to fuse (zip) two streams together.}
Spell, based on the original \ttt{spell} developed by Johnson~\cite{bentley1985spelling}, is another \unix classic:
  after some preprocessing, it makes clever use of \ttt{comm} to report words not in a dictionary.
Shortest-scripts extracts the 15 shortest scripts in the user's \ttt{PATH}, using the \ttt{file} utility and a higher-order \ttt{wc} via \ttt{xargs}~\cite[pg. 7]{taylor2004wicked}.
Diff and Set-diff compare streams via a \ttt{diff} (in \npu, non-parallelizable) and \ttt{comm} (in \pur), respectively.
Sort-sort uses consecutive \pur commands without interleaving them with a \pur that has the effect of significantly condensing the input size (\eg \ttt{uniq}).
The two \ttt{bi-grams} replicate and shift a stream by one entry to calculate bigrams over the input; the \ttt{optimized} version \todo{has a difference}.

\heading{Results}
% TODO \nv{Need to change terminology in plots---files should be (attempted) parallelism, name benchmark, cat should be merge.}
Fig.~\ref{fig:microbenchmarks} presents the speedup gained by \sys as a function of the level of parallelism (2--64$\times$).
% we experiment with a $200\times$ level of parallelism to see potential overheads

In Tab.~\ref{tab:eval} contains two benchmarks that are not shown in Fig.~\ref{fig:microbenchmarks}, as they are artificial modifications intended to highlight \sys-internal characteristics and not lead to a speedup.
  \todo{highlight 2--3 results on things that were not shown in Fig.7}
%   size grows due to the resulting complexity of coordinating among parallel executions---multiple FIFOs per pipeline stage, merging and splitting, encoding of divide-and-conquer synchronization \etc
% The first number alone is interesting, as it captures \sys's output for the sequential script.
% As there is no parallelism involved, it highlights the initial (bare minimum) cost corresponding to using \sys.
% % this is the size of 
% This change in size does not necessarily translate to observed runtime, as most of the size increase comes from the setup and teardown of FIFOs responsible for interprocess (IPC) communication.
% FIFOs are simply pipes made explicit;
%   the performance characteristics, including in-kernel buffering and synchronization mechanisms, are identical between the two.

\heading{Discussion}
\sys achieves near-linear speedup on low-parallelism configurations.
This is due to a high degree (deep trees) of task-based parallelism from \sys's multiple constructs---\eg \emph{fold}s and \ttt{relay}s---evident in all plots that contain \pur stages.
For example, Sort in $8\times$ spawns 37 nodes:
  8 \ttt{tr} nodes, 8 \ttt{sort} nodes, 7 \ttt{sort -m} nodes, and 14 \ttt{relay} nodes.
In effect, \sys hits the maximum number of parallel processes (and thus achieves optimal performance, \ie one without unnecessary process spawns and context switches) 
significantly earlier than the purported parallelism---$16-32\times$ for a 64-core system.
Asking for parallelism beyond that point degrades speedup, as overheads still increase but no further parallelism is available.
Sort-sort with $32\times$ has 314 nodes (sys: 54m) and with $64\times$ has 634 nodes (sys: 90m).
% 
%  On the other hand opt-bigrams 64 has 255 nodes and the sys time stays
%   the same so I think that it might have to do with the fact that
%   constant costs rise a lot (since opt-bigrams 32 takes just 1
%   minute).

@@

\heading{Discussion}
Most pipelines see significant acceleration, except 13, 24, 25, 26, 29, 30 that see no speedup and 2, 19, 31 that see a slowdown.
The first group contain commands which \sys cannot parallelize---\eg \ttt{awk}, \ttt{sed} \ttt{-d}, and \ttt{tr} \ttt{-d}.
Some of them use general interpreters when they could be using a specialized equivalent tailored to the task at hand, the typical example being \ttt{awk} vs \ttt{cut}.
By applying only the replacements which to us experts felt more natural, we also give \sys the opportunity to identify and exploit parallelism---a common pattern we witnessed~\sx{macro1}.
For example, after this change, \todo{13} sees a \todo{XX}$\times$ speedup.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=0.80\columnwidth]{\detokenize{./figs/unix50_individual_speedups_16.pdf}}
  \caption{
	  \textbf{Unix50 scripts.}
    Speedup (left axis) over sequential execution (right axis) for Unix50 scripts.
		Parallelism is 16$\times$ on 10GB of input data~\cf{unix50}.
  }
  \vspace{-15pt}
  \label{fig:unix50-individual}
\end{figure}

@@

\heading{Program}
Coarsely, the pipeline is comprised of two distinct components % coarsely corresponding to a front- and a back-end
  communicating via \unix pipes.
The first downloads pages and extracts their URLs from their source, feeding back \todo{to} the start of the pipeline;
 the second extracts the text from HTML and applies natural-language processing---\eg trigrams, character conversion, term frequencies---to index it.
Collectively, they contains a total of 52 stages (first: 18; second: 34) written in multiple languages---\eg
its \ttt{url-extract}ion utility is written in JavaScript whereas its \ttt{word-stem}ming utility is in Python.
\sys can still operate on them as their parallelizability properties---\sta for \ttt{url-extract} and \ttt{word-stem}---can be trivially described by annotations.

\heading{Results}
% To simplify execution (and avoid Wikipedia), we have saved a recent Wikipedia version on a server within the network and to redirect all \ttt{curl} requests there.
\sys can operate on the entire program, but as its complexity makes end-to-end measurements difficult we report on the two components components individually.
With parallelism set to \todo{2/16} the execution time achieved by \sys is \todo{X} for the front-end and \todo{Y} for the back-end, while the original execution time is \todo{X} for the front-end and \todo{Y'} for the back-end.

@@

\heading{Discussion}
Since the front-end contains a feedback cycle, \sys \todo{doesn't
  parallelize it}, and in both cases it takes \todo{X seconds} to
complete. However, \sys parallelizes the back-end and achieves a
speedup of \todo{X} with parallelism set to \todo{2/16}. Several
stages of the pipeline are \sta (such as the \ttt{word-stem} script
and the text extraction) and that allows \sys to achieve benefits by
exposing data parallelism. Note that the back-end part of the script
contains 34 stages executing in parallel, so the sequential version
\todo{already benefits from task-based parallelism} \kk{not totally
  true}.

>>>>

\begin{table*}[t]
\center
\footnotesize
% \setlength\tabcolsep{3pt}
\caption{
  \footnotesize{
    \textbf{Summary of shell one-liners}.
    Small pipelines (\ie ones usually developed on-the-fly) drawn from various sources and applied to large datasets.
    % Names correspond to the speedup plots (Fig.~\ref{fig:microbenchmarks}).
    Structure summarizes the different classes of commands used in the script.
    Input and seq. time report on the input size fed to the script and the timing of its sequential execution.
    Nodes and compile time report on \sys's resulting DFG size and compilation time for two indicative parallelization configurations.
    % Highlights notes a few key features of each script.
  }
}
\input{microbenchmarks-table}
\label{tab:eval}
\end{table*}

@@

\heading{Programs}
%>\kk{How can we include a result for grep-light?  This is the IO
%>  intensive CPU-light grep. \sys has no speedup but also no slowdown
%>  (even if it IO heavy). Should we just mention that in text in
%>  passing?}
%>
%>\kk{Remember to mention that for spell and bigrams we dont show other
%>  parallel because they are all close to 1 (no speedup).}
Tab.~\ref{tab:eval} summarizes the collection of programs shown in Fig.~\ref{fig:microbenchmarks}, which in turn reports on their performance characteristics.
Grep and Grep-light are in \sta, with the former centered around an expensive DFA-based backtracking expression.
% a DFA-based regular expression matching phase;
 % while the \unix \ttt{grep} command defaults to a Thompson NFA, this particular expression makes use of DFA-based backtracking patterns resulting in high runtime overhead.
Sort is a short script centered around a \pur command.
Wf and Top-n are based on McIlroy's classic word-counting program~\cite{bentley1986literate};
  they use sorting, rather than tabulation, to identify high-frequency terms in a corpus.
% \tr{The \ttt{bi-gram} pipeline calculates n-grams of base two;
%   it makes clever use of \ttt{tail} to shift a stream by one word and \ttt{paste} to fuse (zip) two streams together.}
Spell, based on the original \ttt{spell} developed by Johnson~\cite{bentley1985spelling}, is another \unix classic:
  after some preprocessing, it makes clever use of \ttt{comm} to report words not in a dictionary.
Shortest-scripts extracts the 15 shortest scripts in the user's \ttt{PATH}, using the \ttt{file} utility and a higher-order \ttt{wc} via \ttt{xargs}~\cite[pg. 7]{taylor2004wicked}.
Diff and Set-diff compare streams via a \ttt{diff} (in \npu, non-parallelizable) and \ttt{comm} (in \pur), respectively.
Sort-sort uses consecutive \pur commands without interleaving them with a \pur that has the effect of significantly condensing the input size (\eg \ttt{uniq}).
Finally, the two \ttt{bi-grams} replicate and shift a stream by one entry to calculate bigrams over the input; \ttt{bi-grams-opt} is an optimized version where shift, \ttt{sort}, \ttt{uniq}, have been merged to one command to use a more efficient aggregator.

\heading{Results}
% TODO \nv{Need to change terminology in plots---files should be (attempted) parallelism, name benchmark, cat should be merge.}
Fig.~\ref{fig:microbenchmarks} presents the speedup gained by \sys as a function of the level of parallelism (2--64$\times$). \todo{The average speedups of all one-liners per level of parallelism are 1.97, 3.5, 5.78, 8.83, 10.96, 13.47, and the geometric means are 1.92, 3.39, 5.48, 7.94, 8.86, 8.51.}

% we experiment with a $200\times$ level of parallelism to see potential overheads

Tab.~\ref{tab:eval} contains grep-light, a benchmarks that is not shown in Fig.~\ref{fig:microbenchmarks} since it was designed to show that \sys doesn't slow down IO-intensive scripts. On the contrary, \sys achieves a 1.5-2.5$\times$ speedup for grep-light. In addition, \Cref{tab:eval} shows that \sys compilation times are negligible ($<5s$) for all one-liners.
%   size grows due to the resulting complexity of coordinating among parallel executions---multiple FIFOs per pipeline stage, merging and splitting, encoding of divide-and-conquer synchronization \etc
% The first number alone is interesting, as it captures \sys's output for the sequential script.
% As there is no parallelism involved, it highlights the initial (bare minimum) cost corresponding to using \sys.
% % this is the size of 
% This change in size does not necessarily translate to observed runtime, as most of the size increase comes from the setup and teardown of FIFOs responsible for interprocess (IPC) communication.
% FIFOs are simply pipes made explicit;
%   the performance characteristics, including in-kernel buffering and synchronization mechanisms, are identical between the two.

\heading{Discussion}
\sys achieves near-linear speedup on low-parallelism configurations.
This is due to a high degree (deep trees) of task-based parallelism from \sys's multiple constructs---\eg \emph{fold}s and \ttt{relay}s---evident in all plots that contain \pur stages.
For example, Sort in $8\times$ spawns 37 nodes:
  8 \ttt{tr} nodes, 8 \ttt{sort} nodes, 7 \ttt{sort -m} nodes, and 14 \ttt{relay} nodes.
In effect, \sys hits the maximum number of parallel processes (and thus achieves optimal performance, \ie one without unnecessary process spawns and context switches) 
significantly earlier than the purported parallelism---$16-32\times$ for a 64-core system.
Asking for parallelism beyond that point degrades speedup, as overheads still increase but no further parallelism is available.
Sort-sort with $32\times$ has 314 nodes \todo{(sys: 54m)} and with $64\times$ has 634 nodes \todo{(sys: 90m)}.
% 
%  On the other hand opt-bigrams 64 has 255 nodes and the sys time stays
%   the same so I think that it might have to do with the fact that
%   constant costs rise a lot (since opt-bigrams 32 takes just 1
%   minute).

@@

\heading{Discussion}
Most pipelines see significant acceleration, except 13, 24, 25, 26, 29, 30 that see no speedup and 2, 19, 31 that see a slowdown.
The first group contain commands which \sys cannot parallelize---\eg \ttt{awk}, \ttt{sed} \ttt{-d}, and \ttt{tr} \ttt{-d}.
Some of them use general interpreters when they could be using a specialized equivalent tailored to the task at hand, the typical example being \ttt{awk} vs \ttt{cut}.
By applying only the replacements which to us experts felt more natural, we also give \sys the opportunity to identify and exploit parallelism---a common pattern we witnessed~\sx{macro1}.
For example, after this change, 13 sees a 8.1$\times$ speedup.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{\detokenize{./figs/unix50_individual_speedups_16.pdf}}
  \caption{
	  \textbf{Unix50 scripts.}
    Speedup (left axis) over sequential execution (right axis) for Unix50 scripts.
		Parallelism is 16$\times$ on 10GB of input data~\cf{unix50}.
  }
  \vspace{-15pt}
  \label{fig:unix50-individual}
\end{figure}

@@

\heading{Program}
Coarsely, the pipeline is comprised of two distinct components % coarsely corresponding to a front- and a back-end
  communicating via \unix pipes.
The first downloads pages and extracts their URLs from their source, feeding back \todo{to} the start of the pipeline;
 the second extracts the text from HTML and applies natural-language processing---\eg trigrams, character conversion, term frequencies---to index it.
Collectively, they contains a total of 52 stages (first: 18; second: 34) written in multiple languages---\eg
its \ttt{url-extract}ion utility is written in JavaScript whereas its \ttt{word-stem}ming utility is in Python.
\sys can still operate on them as their parallelizability properties---\sta for \ttt{url-extract} and \ttt{word-stem}---can be trivially described by annotations. \kk{Mention the input. 1/100 of the whole Wikipedia downloaded in a local directory.}

\heading{Results}
% To simplify execution (and avoid Wikipedia), we have saved a recent Wikipedia version on a server within the network and to redirect all \ttt{curl} requests there.
\sys can operate on the entire program, but as its complexity makes end-to-end measurements difficult we report on the two components components individually.
With parallelism set to 2/16 the execution time achieved by \sys is \todo{X} for the front-end and 97m/15m for the back-end, while the original execution time is \todo{X} for the front-end and 191m for the back-end.

@@

\heading{Discussion}
Since the front-end contains a feedback cycle, \sys \todo{doesn't
  parallelize it}, and in both cases it takes \todo{X seconds} to
complete. However, \sys parallelizes the back-end and achieves a
speedup of 1.97/12.7x with parallelism set to 2/16. Several
stages of the pipeline are \sta (such as the \ttt{word-stem} script
and the text extraction) and that allows \sys to achieve benefits by
exposing data parallelism. Note that the back-end part of the script
contains 34 stages executing in parallel, so the sequential version
already benefits from task-based parallelism.
